A Statistical Dynamical Model to Predict Extreme Events and Anomalous Features in Shallow Water Waves with Abrupt Depth Change

### 2018-11-27 paper1_tKdV/tKdV_theory.tex
NM

\begin{figure*}[h]
\centering
\subfloat[downstream state $D_{0}=0.24$]{\includegraphics[scale=.95]{./fig3_1}}
\vspace{-1em}
\subfloat[upstream state $D_{0}=1$]{\includegraphics[scale=.95]{./fig3_2}}
\vspace{-1em}
\caption{Realization of the downstream and upstream flow solutions $u_{\Lambda}^{\pm}$. Note the larger vertical scale in the downstream time-series plot.\label{fig:Realization}}
\end{figure*}
\begin{itemize}
\item The transition from near-Gaussian to skewed non-Gaussian distribution
as well as the jump in both skewness and kurtosis observed in the
experiment observations (Fig. 1 of  \cite{bolles2018anomalous}) can
be characterized by the statistical model simulation results (see
the first and second row of Figure \ref{fig:Changes-stats}). Notice
that the difference in the decay of third and fourth moments in the far
end of the downstream regime from the experimental data is due to the
dissipation effect in the flow from the wave absorbers that is not modeled in the statistical
model here. The model simulation time-series plotted in Figure \ref{fig:Realization}
can be compared with the observed time sequences from experiments
(Fig. 1 of \cite{bolles2018anomalous}). The downstream simulation generates
waves with strong and frequent intermittency towards the positive
displacement, while the upstream waves show symmetric displacements
in two directions with at most small peaks in slow time. Even in the
time-series at a single location $x=0$, the long-time variation displays similar
structures. 
\item The downstream PDFs in experimental data are estimated with a Gamma
distribution in Fig. 2 of \cite{bolles2018anomalous}. Here in the same
way, we can fit the highly skewed outgoing flow PDFs from the numerical
results with the Gamma distribution
\[
\rho\left(u;k,\alpha\right)=\frac{e^{-k}\alpha^{-1}}{\Gamma\left(k\right)}\left(k+\alpha^{-1}u\right)^{k-1}e^{-\alpha^{-1}u}.
\]
The parameters $\left(k,\alpha\right)$ in the Gamma distribution
are fitted according to the measured statistics in skewness and variance,
that is, $\sigma^{2}=k\alpha^{2},\kappa_{3}=2/\sqrt{k}$. And the excess
kurtosis of the Gamma distribution can be recovered as $\kappa_{4}=6/k$.
As shown in the third row of Figure \ref{fig:Changes-stats}, excellent
agreement in the PDFs with the Gamma distributions
is reached in consistency with the experimental data observations.
The accuracy with this approximation increases as the initial inverse
temperature $\theta^{-}$ increases in value to generate more skewed
distribution functions.
\item The experiments also have the up and down stream power spectra in
time (Fig. 4 of \cite{bolles2018anomalous}), which shows more energy
at small time scales, i.e., a relatively slower decay rate in the
downstream compared with the upstream case. This is also observed
in the direct numerical simulations here (detailed results shown
in \textcolor{blue}{\emph{SI Appendix, C.2}}). The downstream state contains
more energetic high frequencies.
\textcolor{blue}{The peak frequency illustrates the characteristic time scale of the transporting
wave trains along the water tank}.

\todo[inline]{I do not understand the last sentence above. Why does the peak frequency illustrate the occurrence of transporting water waves?}

>>>>

\begin{figure*}[h]
\centering
\subfloat[downstream state $D_{0}=0.24$]{\includegraphics[scale=.95]{./fig3_1}}
\vspace{-1em}
\subfloat[upstream state $D_{0}=1$]{\includegraphics[scale=.95]{./fig3_2}}
\vspace{-1em}
\caption{Realization of the downstream and upstream flow solutions $u_{\Lambda}^{\pm}$. Note the larger vertical scale in the downstream time-series plot.\label{fig:Realization}}
\end{figure*}
\begin{itemize}
\item The transition from near-Gaussian to skewed non-Gaussian distribution
as well as the jump in both skewness and kurtosis observed in the
experiment observations (Fig. 1 of  \cite{bolles2018anomalous}) can
be characterized by the statistical model simulation results (see
the first and second row of Figure \ref{fig:Changes-stats}). Notice
that the difference in the decay of third and fourth moments in the far
end of the downstream regime from the experimental data is due to the
dissipation effect in the flow from the wave absorbers that is not modeled in the statistical
model here. The model simulation time-series plotted in Figure \ref{fig:Realization}
can be compared with the observed time sequences from experiments
(Fig. 1 of \cite{bolles2018anomalous}). The downstream simulation generates
waves with strong and frequent intermittency towards the positive
displacement, while the upstream waves show symmetric displacements
in two directions with at most small peaks in slow time. Even in the
time-series at a single location $x=0$, the long-time variation displays similar
structures. 
\item The downstream PDFs in experimental data are estimated with a Gamma
distribution in Fig. 2 of \cite{bolles2018anomalous}. Here in the same
way, we can fit the highly skewed outgoing flow PDFs from the numerical
results with the Gamma distribution
\[
\rho\left(u;k,\alpha\right)=\frac{e^{-k}\alpha^{-1}}{\Gamma\left(k\right)}\left(k+\alpha^{-1}u\right)^{k-1}e^{-\alpha^{-1}u}.
\]
The parameters $\left(k,\alpha\right)$ in the Gamma distribution
are fitted according to the measured statistics in skewness and variance,
that is, $\sigma^{2}=k\alpha^{2},\kappa_{3}=2/\sqrt{k}$. And the excess
kurtosis of the Gamma distribution can be recovered as $\kappa_{4}=6/k$.
As shown in the third row of Figure \ref{fig:Changes-stats}, excellent
agreement in the PDFs with the Gamma distributions
is reached in consistency with the experimental data observations.
The accuracy with this approximation increases as the initial inverse
temperature $\theta^{-}$ increases in value to generate more skewed
distribution functions.
\item Experimental measurements of the power spectra (Fig. 4 of \cite{bolles2018anomalous}) reveal the downstream measurements to contain more energy at small scales, i.e. a relatively slower decay rate of the spectrum. 
This result is also observed
in the direct numerical simulations here (detailed results shown
in \textcolor{blue}{\emph{SI Appendix, C.2}}), as the outgoing state contains
more energetic high frequencies.
